---
title: "Bid caching report"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
source('session1.R')
```
Source
``` {r filename}
cat(filenames[filenum])
```

Number of impressions in sample:
```{r imps}
cat(bid.cache.summary$impressions.in.sample)
```
impression with non-blank cookie
```{r nonblank}
cat(sep="",bid.cache.summary$impressions.with.uid, ' (', round(100*bid.cache.summary$impressions.with.uid/bid.cache.summary$impressions.in.sample),'%)')
```
Number of follow-on impressions from same cookie within 60 seconds
```{r sec60}
cat(sep="",bid.cache.summary$impressions.with.perceding.within.time, ' (', 
    round(100*bid.cache.summary$impressions.with.perceding.within.time/ bid.cache.summary$impressions.with.uid),'%)') 
```
Number of impressions where bid can be improved
``` {r impressions.added.value}
cat(sep="",bid.cache.summary$impressions.with.win.added.value, ' (',
  round(100*bid.cache.summary$impressions.with.win.added.value/ bid.cache.summary$impressions.with.perceding.within.time),'%)')
```
Number and value of reused bids that were late in original auction
``` {r reused_late}
cat(sep="",
    bid.cache.summary$impressions.where.added.value.is.from.late.bids  ,
    ' (',
    round(100*bid.cache.summary$impressions.where.added.value.is.from.late.bids/bid.cache.summary$impressions.with.win.added.value), 
    '%), $',
    round(bid.cache.summary$added.value.from.late.bids/10)/100) 
```
Number and value of reused bids that were actually served in original auction
``` {r reused_already_served}
cat(sep="",
    bid.cache.summary$impressions.reusing.winner,
    ' (',
    round(100*bid.cache.summary$impressions.reusing.winner/bid.cache.summary$impressions.with.win.added.value), '%, $', round(bid.cache.summary$impressions.reusing.winner.value/10)/100,
    ')'
    )

```

Added bid value
``` {r total.value}
cat(sep="",'$',round(bid.cache.summary$added.bid.value/10)/100, " over base value of $", round(bid.cache.summary$sent.bids.value/10)/100, 
    " (+", round(100*bid.cache.summary$added.bid.value/bid.cache.summary$sent.bids.value), '%)')
```

Improvement Histogram

``` {r improve.hist}
print(p2)
```